WORCS: A Workflow for Open Reproducible Code in Science

### 2020-11-08 README.Rmd
Incorporate Loek's comments

The `worcs` package is open to contributions from collaborators with valuable expertise, to improve both the workflow and the software. Relevant contributions warrant coauthorship to the package. Please contact the lead author at [c.j.vanlissa@uu.nl](mailto:c.j.vanlissa@uu.nl), or:

  - File a GitHub issue [here](https://github.com/cjvanlissa/worcs)
  - Make a pull request
    [here](https://github.com/cjvanlissa/worcs/pulls)

>>>>

We are always eager to receive user feedback and contributions to help us improve both the workflow and the software. Major contributions warrant coauthorship to the package. Please contact the lead author at [c.j.vanlissa@uu.nl](mailto:c.j.vanlissa@uu.nl), or:

  - [File a GitHub issue](https://github.com/cjvanlissa/worcs) for feedback, bug reports or feature requests
  - [Make a pull request](https://github.com/cjvanlissa/worcs/pulls) to contribute your code or prose

### 2020-11-08 paper/Manuscript_DS.Rmd
Incorporate Loek's comments

An integral part of WORCS is the *explicit* version control software Git
(www.git-scm.com).
Git tracks changes to files on a line-by-line basis.
The user can sture these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Git retains a complete log of all commits, and users can compare changes between
different commits, or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's private computer.

>>>>

An integral part of WORCS is the *explicit* version control software Git
(www.git-scm.com).
Git tracks changes to files on a line-by-line basis.
The user can sture these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Each commit is further tagged with a message, describing the changes.
The `R` implementation of `worcs` contains a convenience function,
`git_update("your commit message")`,
which creates a new commit for all changes since the previous commit,
labels it with "your commit message", and pushes these changes to the remote repository.
It is, effectively, a Git "quick save" command.

From an open science perspective, Git is particularly appealing because it
retains a complete historical backlog of all commits.
This can be used, for example, to verify that authors executed the steps outlined in a preregistration.
Users can also compare changes between different commits,
or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's private computer.

### 2020-11-08 paper/reviews_ds.Rmd
Update
Incorporate Loek's comments

Response: We appreciate the recommendation of performing a user evaluation, and this is certainly a next step in the development of worcs. However, as the Editor has indicated, 'resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse"'. We will thus first address the potential for re-use, and then we will address steps we have taken and intend to take with regard to user evaluation.

First, there are several arguments to substantiate the potential for reuse of worcs:

1. WORCS is implemented in an R-package
2. This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network")
3. The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices)
4. The R-package uses only tools that are fully integrated within the RStudio IDE, thus enhancing ease of use
5. The R-package is compatible with other reproducibility solutions, including Docker, drake, and make

@@

1. The preprint has been downloaded 561 times, since being published on 31-05-2020
1. The `worcs` R-package has been downloaded 2401 times from CRAN, since being published on 18-05-2020
1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
1. The video tutorial for worcs has been viewed 362 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
1. WORCS is used to teach reproducibility at Utrecht University
1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
1. The lead author has applied `worcs` in several research projects:
	* https://github.com/cjvanlissa/storm_meta_father_mother
	* https://github.com/cjvanlissa/veni_sysrev
	* https://github.com/cjvanlissa/placebo_adhd
	* https://github.com/cjvanlissa/schumpe
	* https://github.com/cjvanlissa/coping_covid

Third, although we have not conducted a user evaluation, there is some evidence that `worcs` addresses users' needs:

1. Evaluate worcshops!!
2. The issues page on the worcs GitHub profile demonstrates that we have been very responsive in addressing users' feedback (as required by the aforementioned CII Best Practices):
* https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed

@@

\Assignment{CJ, BV, AL}
\WorkInProgress
\Hard

@@

\Assignment{CJ}
\WorkInProgress
\Easy

@@

We agree completely that it is important to make open data findable beyond just providing it freely.
It is a pity that the Reviewer does not provide arguments to support a preference for Zenodo or FigShare over GitHub, because if there is a strong case to be made, we would be keen to include this in the paper.
Without such arguments, as far as we are concerned, these repositories are largely interchangeable.

@@

@@

Response: It does not. It seems to us that this is outside of the scope of the present paper.
If a compelling case could be made by the Reviewer to include such a service,
we are open to the suggestion - but we would need to take on additional coauthors with the required expertise.
<!--What would this entail? Is it within scope?-->

@@

Third, there seems to be some misunderstanding of the workings of existing solutions.
For example, the Reviewer remarks that 
"git capture changes at the level of the line of code/text, whereas a researcher needs an abstraction that informs him/her on the coarse-grained elements [...] so as to have a clear idea of where s/he is within the process and help him/her make the decision of the next steps to undertake".
This comment suggests a fundamental misunderstanding of how Git works.
Although Git tracks file changes on a line-by-line basis,
any number of such changes can be committed together, thereby capturing the more "course-grained elements".
For example, in our response to these reviews, we have created one Git commit for each Reviewer comment, and these commits contain changes to the action letter, manuscript, and vignettes.

The Reviewer's comment does illustrate the importance of clarifying this point in the manuscript as well
To address this comment, the revised manuscript now reads:

> Git tracks changes to files on a line-by-line basis. The user can sture these changes by making a "commit": a snapshot of the version controlled files. Each "commit" can contain as many changes as desired; for example, a whole new paragraph, or many small changes made to address a single reviewer comment.

>>>>

Response: We appreciate the recommendation of performing a user evaluation, and this is certainly a next step in the development of WORCS. At this point, we have written this manuscript detailing the conceptual workflow, and released a software implementation for R-users. The next step, which we are now entering, is to build a user base, and then mount a user evaluation.

To address the Reviewer's comment, we have set up an evaluation questionnaire, and sent it to the attendees of our previous workshops on WORCS. At the time of writing, however, this has not yielded actionable points of improvement.

We do have a strong history of addressing user comments submitted through the GitHub issues page.
WORCS adheres to the Core Infrastructure Initiative "best practices" (CII Best Practices),
which requires that
the channels through which users can provide feedback are clearly indicated,
as they are [here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
It is also evident from the [resolved tickets on GitHub](https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed) that we have been very responsive in actively incorporating user feedback - as also required by the CII Best Practices.

We are committed to continue evaluating user experience, both actively through evaluation forms, and passively through the GitHub issues pages.
We hope that this commitment, along with the evidence of our prior responsiveness to user feedback presented above, assuage the Reviewer's concerns.

Finally, we wish to point out that the Editor has indicated that 'resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse"'. We therefore elaborate on the efforts we have made to ensure that WORCS is reusable, followed by evidence of its current adoption rates.

We have made the following efforts to ensure potential for reuse:

1. WORCS is implemented in an R-package, and thus readily available to R-users
2. This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network"), and can thus be installed through default channels
3. The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices), including clearly indicated channels for user feedback, and being responsive to that feedback
4. The R-package uses only tools that are fully integrated within the RStudio IDE, thus enhancing ease of use
5. The R-package is compatible with other reproducibility solutions, including Docker, drake, and make

@@

1. [The preprint](https://psyarxiv.com/k4wde/) has been downloaded 568 times, since being published on 31-05-2020
1. The `worcs` R-package has been downloaded 2474 times from CRAN, since being published on 18-05-2020
1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
1. The video tutorial for worcs has been viewed 362 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
1. WORCS is used to teach reproducibility at Utrecht University
1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
1. The lead author has applied `worcs` in several research projects:
	* https://github.com/cjvanlissa/storm_meta_father_mother
	* https://github.com/cjvanlissa/veni_sysrev
	* https://github.com/cjvanlissa/placebo_adhd
	* https://github.com/cjvanlissa/schumpe
	* https://github.com/cjvanlissa/coping_covid

@@

\Assignment{CJ, BV, AL}
\Done
\Hard

@@

@@

We agree completely that it is important to make open data findable beyond just providing it freely.
This is achieved by linking the GitHub repository to an OSF.io project page;
on OSF, the project receives a DOI.
It is also possible to give a separate DOI to specific resources, such as data files.
We agree that this was not emphasized enough in the manuscript, and have rewritten the revision to clarify.

It is a pity that the Reviewer does not provide arguments to support a preference for Zenodo or FigShare over GitHub, because if there is a strong case to be made, we would be keen to include this in the paper.
Without such arguments, as far as we are concerned, these repositories are largely interchangeable.

@@

Response: 
F: Dit heeft weer met DOIs te maken. Ik denk dat als we het gebruik van OSF de default van worcs maken (die mijn opmerking R1C7a) dat we hier een sterk punt kunnen maken mbt findability. 
A: Eigenlijk hetzelfde antwoord als bij Findability: dit wordt heel eenvouding als we de OSF route voor data storage de default maken. 
I: Het feit dat we alleen gebruik maken van open-source software maakt alle projecten die worcs gebruiken nogal interoperable. 
R: Dit is wel een belangrijkt punt: Kunnen we opnemen in de workflow dat worcs-projecten by default een CC-BY licentie krijgen?

@@

Response: WORCS does not provide a service for de-identification,
and we consider this to be outside of the scope of the present paper.
If the Reviewer feels differently, we are open to being convinced, however!
Of course, since `worcs` is open source software, anybody is welcome to contribute this functionality in exchange for coauthorship, in line with the contribution guidelines [outlined here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
The current team lacks relevant expertise to make this contribution.
<!--What would this entail? Is it within scope?-->

\Assignment{ALL}
\Done
\Medium

@@

Third, there seems to be some misunderstanding of the workings of existing solutions.
For example, the Reviewer remarks that 
"git capture changes at the level of the line of code/text, whereas a researcher needs an abstraction that informs him/her on the coarse-grained elements [...] so as to have a clear idea of where s/he is within the process and help him/her make the decision of the next steps to undertake".
This comment suggests a fundamental misunderstanding of how Git works.
Although Git tracks file changes on a line-by-line basis,
such changes are typically grouped together in a "commit",
which is how Git captures the more "course-grained elements" the Reviewer refers to.
For example, in our response to these reviews, we have created one Git commit for each Reviewer comment, and these commits contain changes to the action letter, manuscript, and vignettes.

To address the Reviewer's comment, we now clarify this point in the revised manuscript:

> Git tracks changes to files on a line-by-line basis. The user can sture these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Each commit is further tagged with a message, describing the changes.
The `R` implementation of `worcs` contains a convenience function,
`git_update("your commit message")`,
which creates a new commit for all changes since the previous commit,
labels it with "your commit message", and pushes these changes to the remote repository.
It is, effectively, a Git "quick save" command.
